Efficient and Robust Plane-Based Boolean Operations on N Meshes

### 2017-01-06 paper.tex
change to split files

\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\floatname{algorithm}{Procedure}

\begin{document}
\title{Efficient Non-incremental Constructive Solid Geometry Evaluation for Triangular Meshes}
\author{Rui~Wang,~
        Xudong~Jiang,~
        Hongbo~Fu,~
        Bin~Sheng,~and
        Enhua~Wu
\IEEEcompsocitemizethanks{

@@

\IEEEtitleabstractindextext{
\begin{abstract}
  In this paper, we propose an efficient non-incremental approach to evaluate the boundary of constructive solid geometry (CSG). The face membership classification step is a bottleneck in many existing CSG evaluation approaches. The key idea of our approach is to use local coherence of space labels to accelerate this step. We employ a two-level grouping scheme to group faces that share the same space labels. Therefore, these common space labels can be shared within each group to avoid repetitive computation. Our method is optimized for a non-incremental evaluation. It generates the entire result model at once, rather than performing a step-by-step evaluation of Boolean operations. To achieve a robust evaluation, our approach introduces plane-based geometry into the intersection computation step. Comparison experiments with state-of-the-art techniques show that our method can more efficiently perform boundary evaluation of both trivial and complicated CSG with massive faces while maintaining robustness.
\end{abstract}

\begin{IEEEkeywords}
Boolean operations, triangle mesh, CSG evaluation, plane-based geometry.
\end{IEEEkeywords}

}

\maketitle

\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{C}{onstructive} solid geometry (CSG) has long been a popular modeling tool of computer-aided design and computer-aided manufacturing (CAD/CAM). It constructs complex models by combining primitives using a series of regularized Boolean operations \cite{requicha1977mathematical}: union, intersection and difference. CSG is often represented by a binary tree, called a CSG tree, whose leaves are primitives and whose internal nodes are Boolean operations..

CSG can be converted into the widely-used  boundary representation (i.e., triangle mesh) through boundary evaluation. Most boundary evaluation methods are based first on intersection computation and then on face membership classification. Robustness and efficiency are their major issues. During the last three decades, many techniques have been developed to solve these problems. However, the efficiency of face membership classification still needs improvement.

Computing face space labels is the basis of face classification with respect to CSG. For each face, the number of space labels to be computed is equal to the number of primitives. For large CSG with massive faces and primitives, computing these space labels can be very time-intensive. A common acceleration method is to use localized schemes. This method is based on the local coherence of space labels: if a face is inside (or outside) a specific primitive, the face's neighborhoods are likewise often inside (or outside) the primitive. Determining whether two adjacent faces share the same space labels is relatively simple (see also Section \ref{sec:framework}); therefore, we can locally group together faces of the same labels and reuse space labels to save computation.

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.5in]{cluster}\\
  \caption{2D illustration of grouping schemes. The boundaries of polygons are edges. Suppose each input mesh consists of hundreds of edges (not shown in the figure). Edges are grouped using (a) voxels, (b) octree cells, and (c) edges as the grouping units. Different groups in (c) are marked by different colors.} \label{fig:cluster}
\end{figure}

Several recent studies \cite{campen2010exact, pavic2010hybrid, wang2011approximate} have used such localized schemes. These approaches often use the cube (e.g., voxels, octree cells) as the basic grouping unit to reuse the space division data structures constructed during the intersection computation. However, using the cube as a basic grouping unit has several disadvantages. In general, connected faces with the same space labels should be grouped together. Thus, an ideal face group is a union of connected faces and can be of any shape. The cube-based grouping scheme can only provide a rough approximation of that shape. Therefore, extra computation, which is usually time-consuming, is needed to classify those ungrouped faces. For example, in Figs. \ref{fig:cluster}(a) and (b), the red cubes contain the intersection of two primitives. Thus, the edges (faces in 3D) in these red cubes should belong to different groups (because they have different labels). However, the size of cubes cannot be illimitably small to distinguish these different groups. Therefore, these edges in the red cubes are left as ungrouped edges. In contrast, a face-based grouping scheme can easily handle groups of arbitrary shapes and is potentially more efficient for face classification (Fig. \ref{fig:cluster}(c)).

Another limitation of prior exact evaluation algorithms is that most of them are incremental ones \cite{campen2010exact, feito2013fast, bernstein2009fast}. This means that they can evaluate only one Boolean operation at a time. For a large CSG tree with more than two primitives, incremental algorithms have to decompose the CSG into a series of Boolean operations and then incrementally evaluate each one. This strategy is highly inefficient because it unnecessarily computes the CSG intermediate results. Incremental algorithms have been more popular for a considerable time because human designers typically construct CSG models by progressively adding primitives. However, recent techniques---such as GPU-based approximate evaluation algorithms \cite{pavic2010hybrid, wang2011approximate} and CSG visualization algorithms \cite{hable2007cst, rossignac2011ordered}---are more suitable for such a preview computation. Therefore, exact evaluation algorithms are now more likely to be used for final mesh generation (after all the primitives are added), where a non-incremental algorithm has more advantages.

In this paper, we propose a method to perform CSG boundary evaluation with triangular mesh primitives. The proposed method uses the face (i.e., a triangle in our implementation) as the basic unit of the grouping and applies a flood-filling algorithm to group faces. Our method utilizes the local coherence of space labels. It propagates face space labels between adjacent faces and adjacent groups to save computation of space labels. The most prominent characteristics of our method are outlined below.
\vspace{0.0em}
\begin{itemize}[leftmargin=*]
  \item{\textbf{Face classification using two-level grouping}}~~~~To efficiently classify faces, we use a two-level grouping scheme to reuse space labels. Input faces are firstly grouped into first-level groups at the primitive level; then, the first-level groups are further divided into second-level groups according to their labels (Fig. \ref{fig_overview}(d)). A flood-filling algorithm is applied to enable efficient grouping and label sharing among adjacent faces. This framework makes a balance between the benefit of face label sharing and the cost of grouping faces for the best performance.

  \vspace{0.5em}
  \item{\textbf{Efficient non-incremental evaluation}}~~~~Unlike many existing Boolean methods \cite{granados2003boolean, bernstein2009fast, campen2010exact, feito2013fast}, the proposed method can perform non-incremental evaluation of CSG, which computes the entire final model directly. Our approach is more efficient at evaluating large CSGs because it computes no intermediate results in the internal nodes of CSG trees. A non-incremental evaluation is more difficult to perform because the conditions of the intersection and face classification are more complicated. We developed a set of techniques, including face-nested binary space partitioning (BSP) and multi-level CSG trimming, to make our method highly efficient for non-incremental evaluation.

  \vspace{0.5em}
  \item{\textbf{Robustness and exactness}}~~~~In the field of solid modeling, efficiency without robustness is impractical. We therefore adopt a series of schemes to make our method robust. To avoid the introduction of errors during intersection computation, we combine a triangle-triangle intersection algorithm \cite{moller1997fast} with a plane-based representation \cite{sugihara1990solid}. In addition, an exact BSP construction technique \cite{bernstein2009fast, campen2010exact} is adopted to enhance the robustness of face classification.
\end{itemize}
\vspace{0.5em}

Experiments performed confirm that our method is more efficient than state-of-the-art techniques \cite{cgal:hk-bonp3-15a, campen2010exact, feito2013fast, Maya2015}. It can quickly and robustly perform CSG evaluations not only for trivial CSGs, such as one with single Boolean operation, but also for large CSGs with hundreds of primitives.

\section{Related Work}

\begin{figure*}[!t]
\centering
\includegraphics[width=7.1in]{flowchart}
\caption{An overview of the proposed method for CSG evaluation. In this toy example, the CSG tree represents a solid constructed by (Cube $\cup$ Sphere - Cylinder). (d) shows how to classify faces step by step using a flood filling algorithm. Faces of the same primitive are grouped as the inter-primitive group. Then faces of the same inter-primitive group are further divided into intra-primitive groups according to their conditions of intersection (c). The unvisited faces are black. Different colors of other faces indicate that they are accepted and belong to different intra-primitive groups. Rejected faces are rendered with wire-frame. The seed face of flood filling in each primitive is highlighted in red.}
\label{fig_overview}
\end{figure*}

CSG evaluation has had notorious problems with robustness since its inception in the 1980s \cite{requicha1985boolean, laidlaw1986constructive}. The non-robustness is inherited from the building blocks of CSG: Boolean operations on solids. Many researchers have attempted to solve such an issue using arbitrary precision arithmetic \cite{banerjee1996topologically, fortune1995polyhedral, keyser2004esolid, granados2003boolean, hachenberger2005boolean} and exact interval computation \cite{fang1993robustness, hu1996robust, segal1990using}. However, these methods are often too expensive in computation time and memory to be practical for evaluation of CSG with massive faces. For example, in the Computational Geometry Algorithms Library (CGAL) \cite{cgal:hk-bonp3-15a}, the state-of-the-art robust Boolean operation algorithm \cite{hachenberger2005boolean} (implemented with arbitrary precision arithmetic) is more than 20 times slower than its non-robust version.

Sugihara and Iri \cite{sugihara1990solid} introduced a plane-based representation of polyhedra. They found that Boolean operations are fundamentally robust using the plane equation as the primary geometry representation. In this condition, the evaluation of Boolean operations can be performed with no `constructions' \cite{shewchuk1999lecture}, thereby avoiding the introduction of any numerical errors.

Berstein and Fussell \cite{bernstein2009fast} noted that a binary space partitioning (BSP) merging algorithm in Boolean operations \cite{thibault1987set, naylor1990merging} is actually a plane-based technique. Therefore, they combined the two conceptions---plane-based geometry and BSP merging---to develop an unconditionally robust method for Boolean operations of polyhedra. Using Shewchuk's adaptive geometry predicate technique \cite{shewchuk1997adaptive}, this method can be only twice slower than non-robust methods. Our approach adopts this technique in the stage of face classification to ensure the robustness.

Berstein and Fussell's method does not alleviate the problem of high memory consumption of BSP-based algorithms. Thus, it remains unprepared for polyhedra with massive faces. However, it inspired Campen and Kobbelt to develop a more delicate approach  \cite{campen2010exact}. This approach solves the problem of exact and efficient conversion of polyhedra between vertex-based and plane-based representations. Moreover, by leveraging the adaptive octree, BSP structures are nested into octree cells where the intersection between primitives occurs. For efficiency, face classification is performed using cells as the basic classification units. This localized scheme saves considerable computing resources and retains the topology of non-intersected areas.

Such localized scheme is widely used in different methods of CSG boundary evaluation \cite{pavic2010hybrid, wang2011approximate, feito2013fast, campen2010exact}. These methods are usually based on intersection computation and face membership classification. The success of such localized scheme is enabled by two facts. First, the intersections between primitives are locally distributed. Second, space labels of a face are often coherent with those of its neighborhood. The former makes it possible to reduce the space that requires an intersection test. The latter enables sharing of space labels between adjacent faces to accelerate face classification.

However, most of these methods use the cube (e.g., voxels, octree cells) as the unit of the face group, which is often the natural result of reusing space-division data structures constructed for intersection computation. In this way, connected faces that share the same labels might be distributed in different cubes. Thus, repetitive classifications may be required for each cube. Furthermore, faces of different classifications may coexist within a small space, forming a ‘mixed’ area (e.g., the red areas in Figs. \ref{fig:cluster}(a) and (b)). Because the size of the cubes cannot be illimitably small, special cubes that contain faces with different space labels typically must be constructed in these mixed areas. Processing these special cubes is often complex and time-consuming. To more efficiently share space labels, we therefore use faces, rather than cubes, as the basic grouping units. Grouping is efficiently performed by utilizing geometry connectivity with a flood-filling algorithm.

Recently, Feito et al. \cite{feito2013fast} proposed a method for Boolean evaluation. This method also uses a face-based grouping scheme. To share classification results, faces that do not intersect with other primitives are grouped together according to geometry connectivity. This method and the proposed method have similar stages for single Boolean operation evaluation. One of the major differences is that our method supports non-incremental evaluation of CSG. For this purpose, we designed a series of dedicated stages. Additionally, our face grouping scheme is intended for both non-intersected faces and faces that intersect with other primitives. It thus provides a uniform treatment for all faces and further accelerates the face classification stage. Moreover, the similarity between adjacent groups is omitted by Feito et al.'s method. Although different groups have different space labels, we observe that they may still share some of these labels if they are neighboring. We utilize this kind of space label coherence between groups for better performance.

On the other hand, with the development of general-purpose computing on graphics processing units (GPGPUs), many researchers  \cite{wang2011approximate, zhao2011parallel, museth2002level, chen1999volumetic, eisemann2008single} have tried to utilize the grand computation power of graphics hardware for Boolean operations. These methods often have good performance and are suitable for interactive applications, such as digital sculpting. However, owing to the paralleled features of graphics hardware, these methods are usually voxel-based and support only approximate evaluation that inevitably suffers from geometric detail loss, especially at the intersection areas of primitives.

\section{Overview}

\label{sec:overview}

As in most previous works \cite{feito2013fast, bernstein2009fast}, we require the input primitives of the CSG tree to be watertight Nef polyhedra with manifold surfaces. We assume there is no self-intersection. Because we adopt a flood-filling strategy, we additionally assume that the connectivity of adjacent faces is available. Our current implementation is dedicated to primitives of triangular meshes for simplicity. Our approach efficiently and robustly computes the boundary of CSG solids in three steps, as illustrated in Fig. \ref{fig_overview}. Our main contributions exist in Steps 2 and 3.

\subsection{Adaptive Octree Construction}

As a typical routine of localization, our method uses the adaptive octree to frequently accelerate the used spatial query afterwards. Our octree construction implementation is akin to the implementation in Feito et al.'s method \cite{feito2013fast}. The major difference is that, because our method is for non-incremental CSG evaluation, the octree is constructed on the bounding box that contains all input primitives (Fig. \ref{fig_overview}(b)). To avoid presentation ambiguity between a CSG leaf and an octree leaf, the latter is called a cell. Intersection detection between triangle faces and cells can be efficiently performed by using the separating axis theorem \cite{gottschalk1996obbtree}. Cells are classified into two types: if all triangles that intersect a cell belong to the same primitive, we call it a \emph{non-critical} cell. Otherwise, it is a \emph{critical} cell. This classification is needed for intersection computation in the next step.

\subsection{Plane-based Intersection Computation}
\label{sec:overview:isect}

In this step, we compute all intersections between primitives and record them in intersected faces (e.g., the red triangles in Fig. \ref{fig_overview}(c)) for further computation. Because we have already constructed the octree and obtained its cell classification, it is instantly known that intersection between faces occurs only within critical cells. Thus, we limit the intersection test within these critical cells. The intersection test is performed for every pair of faces in each critical cell. We assume that primitives are not self-intersected; therefore, tests for faces from the same primitives are unnecessary.

The triangle-triangle intersection test always incurs robustness issues. To avoid an unexpected failure caused by numeric errors, we integrate a plane-based representation of polyhedra \cite{sugihara1990solid} into the triangle-triangle intersection test. Adaptive precision predicate technique \cite{shewchuk1997adaptive} is applied for efficiency of plane-based geometry computation. Details are provided in Section \ref{section:isect}.

\subsection{Face Classification Using Two-level Grouping}

Face classification using two-level grouping is the most novel step of our algorithm. For each primitive face, we must determine whether it, or its part (when the face intersects with other faces), belongs to the final model. A breadth-first flood-filling algorithm is used to traverse faces of each primitive. Starting from a random seed whose space labels are all computed, these labels propagate to neighboring faces and vary according to intersection conditions (Fig. \ref{fig_overview}(d)). Each face is instantly classified once all of its space labels are computed. We use a two-level face grouping scheme to enable the maximum information reuse. The first level is the \emph{inter-primitive} level; the second is the \emph{intra-primitive} level. When visiting an intersected face, constrained Delaunay triangulation \cite{chew1989constrained} is applied to subdivide the face into smaller triangles, followed by individual classification of the subdivided triangles using BSP techniques. The robustness of this step is ensured by the exact BSP computation technique \cite{bernstein2009fast}. Details are shown in Section \ref{section:floodfill}.

\section{Plane-based Intersection Computation}

\label{section:isect}

In this step, intersections between faces are computed through the triangle-triangle intersection test within each critical cell. We adopt M\"{o}ller's intersection detection algorithm \cite{moller1997fast} for this test on account of the algorithm's efficiency and simplicity. However, because collision detection algorithms are often accompanied by non-robustness problems, a conventional implementation of the algorithm may cause unpredictable results. Therefore, we integrate plane-based geometry representation \cite{sugihara1990solid} into M\"{o}ller's algorithm to make it unconditionally precise and robust.

\subsection{Intersection Detection}

\begin{figure}[t]
\centering
\includegraphics[width=3.0in]{projection}
\caption{$Seg_1$ is the intersection between $S_2$ and $F_1$. $Seg_2$ is the intersection between $S_1$ and $F_2$. The intersection between $F_1$ and $F_2$, which is the line segment in red, is the overlap of $Seg_1$ and $Seg_2$.}
\label{fig_projection}
\end{figure}

M\"{o}ller's algorithm computes the intersection between two triangles $F_1$ and $F_2$ in three steps, as illustrated in Fig. \ref{fig_projection}. First, an early rejection is performed by testing whether $F_1$ intersects the plane where $F_2$ lies, which is denoted as $S_2$. This is a necessary condition of intersection between $F_1$ and $F_2$. If this test is passed, then the same is performed for $F_2$ and $S_1$, where $S_1$ is the plane of $F_1$. Second, the intersection between $F_1$ and $S_2$, denoted as $Seg_1$, and the intersection between $F_2$ and $S_1$, denoted as $Seg_2$, are separately computed . At last, the intersection between $F_1$ and $F_2$ is determined by computing the overlap area of $Seg_1$ and $Seg_2$ .

\begin{figure}[t]
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=3.0in]{nonrobust}\\
  \caption{The left edges of A and B (\emph{left}) are nearly but not exactly collinear. However, under usual float point arithmetic, they might be judged as collinear causing discontinuous edges in the final result (\emph{right}).}\label{fig:precision}
\end{figure}

Conventional implementations of M\"{o}ller's algorithm use vertex-based representation and usual float point arithmetic. However, this is neither exact nor robust. In Fig. \ref{fig:precision}, we illustrate a non-robust 2D case of the Boolean operation. Although implementing M\"{o}ller's algorithm with arbitrary precision arithmetic could make it robust, it is too costly for a large CSG evaluation. Actually, the non-robustness of this algorithm is from \emph{constructions} \cite{bernstein2009fast}, which compute new coordinates of geometry objects based on the known coordinates of existing ones. In M\"{o}ller's algorithm, the coordinates of $Seg_1$ and $Seg_2$ (Fig. \ref{fig_projection}) are computed as intermediate results. On the other hand, if the computation can be restricted to \emph{predicates}, which make a two or three-way decision based on known coordinates, we could quickly and robustly implement this algorithm.

Fortunately, according to Sugihara and Iri \cite{sugihara1990solid}, geometry computation of Boolean operations can be restricted to predicates using plane-based representation of polyhedra. This motivated us to integrate plane-based representation into M\"{o}ller's algorithm. Exact and efficient predicate computation is performed using precise geometry predicate techniques \cite{shewchuk1997adaptive}. Details of our implementation are shown below.

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{p-reps}\\
  \caption{Plane-based representation of triangles. The yellow triangle is represented by three bounding planes (green) and the supporting plane where the triangle lies.}\label{fig:p-reps}
\end{figure}

We first convert $F_1$ and $F_2$ into their respective plane-based representations: a supporting plane surrounded by three bounding planes (Fig. \ref{fig:p-reps}). We implement this conversion using the exact conversion method by Campen and Kobbelt \cite{campen2010exact}. Then, in our implementation, the relative position of a point and a plane can be determined by computing the sign of the multiplication of determinants \cite{sugihara1990solid}. To avoid introducing errors, we represent the end points of $Seg_1$ and $Seg_2$  by intersections of plane triples, rather than by directly computing their coordinates. As illustrated in Fig. \ref{fig:isect}(a), these points are the intersection between an edge line of a face and the supporting plane of the other face. Because the edge line of $F_1$ can be represented as $[S_1, S_X]$ (or $[S_2, S_X]$ for an edge line of $F_2$), where $S_X$ is the corresponding bounding plane of the edge, the end points of $Seg_1$ or $Seg_2$ can be represented in the form of $[S_1,S_2,S_X]$. In Fig. \ref{fig:isect}, we list all intersection situations between a triangle and a plane, and the $S_X$  in each situation.

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{sign}
\caption{We denote the signed distance from point $X$ to plane $S_2$ as $d_X$. All the four conditions of intersection between $F_1$ and $S_2$ (denoted as $Seg_1$) are:  (a) $d_A\cdot d_C<0$, $d_B\cdot d_C<0$; (b) $d_A=0$, $d_B=0$, $d_C\neq 0$; (c) $d_A=0$, $d_B\cdot d_C<0$; (d) $d_A=0$, $d_B\cdot d_C>0$. End points of $Seg_1$ are intersections between $S_2$ and related edge lines of $F_1$ (bold edges).}
\label{fig:isect}
\end{figure}

The intersection between $F_1$ and $F_2$ is the overlap area of $Seg_1$ and $Seg_2$. It can be easily computed by comparing the end points of $Seg_1$ and $Seg_2$ along $L$ (Fig. \ref{fig_projection}), where $L$ is the intersection between $S_1$ and $S_2$. Unlike M\"{o}ller's implementation, our approach uses no projection, which requires the exact point coordinates. We call the direction of vector $\vec{N}_1\times \vec{N}_2$ the positive direction, where $\vec{N}_1$ and $\vec{N}_2$ are normals of $S_1$ and $S_2$ respectively. Given two points $A$ and $B$ on $L$,  we compute the sign of the multiplication of determinants:
\begin{equation}
\label{eq:sign}
  K(A, B)=
  \left|
  \begin{array}{c}   %该矩阵一共3列，每一列都居中放置
  \vec{S}_1 \\ \vec{S}_2 \\ \vec{S}_A \\ \vec{S}_B
  \end{array}
  \right|
  \left|                 %左括号
  \begin{array}{c}   %该矩阵一共3列，每一列都居中放置
  \vec{N}_1 \\\vec{N}_2 \\\vec{N}_A
  \end{array}
  \right|
  \left|                 %左括号
  \begin{array}{c}   %该矩阵一共3列，每一列都居中放置
  \vec{N}_1 \\\vec{N}_2 \\\vec{N}_B
  \end{array}
  \right|,
\end{equation}
where $\vec{S}_X$ is the coefficient vector of plane equation $[a_X,b_X,c_X,d_X]$,  and $ \vec{N}_X$ is the normal vector $[a_X,b_X,c_X]$. If $K(A, B)$ is positive (negative), this means that $A$ lies in the positive (negative) direction of $B$. Otherwise, $A$ and $B$ are concident.

If $F_1$ and $F_2$ are coplanar, we check whether they overlap within the common plane using the algorithm by M\"{o}ller\cite{moller1997fast}. The substrate of this algorithm is to verify on which side of an edge $E$ a point $V$ lies. We implement it using plane-based geometry by substituting $E$ with its corresponding bounding plane $S_E$ and then computing the sign of distance from $V$ to $S_E$.

\subsection{Intersection Recording}
\label{sec:isecthandle}

The information of intersection is stored in the intersected faces where the intersection occurs. Because the intersection is generated by two (or more) faces, both (all) faces are intersected faces and store the same copy of intersection information.

Each intersected face maintains two important lists for further processing: \emph{cross list} and \emph{coplanar list}. The cross list stores the end points of intersections, the associated faces of the intersections, and the primitives to which these associated faces belong. The coplanar list is relatively simple. It only stores coplanar faces and the primitives to which these coplanar faces belong.

Moreover, in order to reconstruct geometry connectivity and to avoid coincident vertices in the final model, coincident points detected during intersection computation are merged. A \emph{point list} for each intersected face $F$ is used for this purpose. The list stores all the points inside $F$ or on the boundary of $F$. When introducing new vertices into the face (e.g., recording the new intersection line into the cross list, the two end points of the intersection line segment are newly introduced vertices), we determine if there exists a coincident point in the point list. Coincident points are merged. In addition, because the points on the edges of $F$ are shared by adjacent faces, these points should be added to the corresponding point lists of both two adjacent faces.

There is a degenerate condition for intersection recording: $F_1$ and $F_2$ may intersect on a single point. We treat this condition as $F_1$ and $F_2$ being non-intersecting but add the intersected point into the point lists of both faces. Another point to notice is that if $F_1$ and $F_2$ are coplanar and overlap, $F_1$ is added to the coplanar list of $F_2$ and vice versa. No new point is introduced under this condition.

\section{Face Classification}
\label{section:floodfill}

In this step, we classify faces one by one to generate the final mesh. The intersection information computed in the previous section is used for re-tessellation and space label computation. To better explain our approach, we firstly introduce the background knowledge. Then the framework of this step is outlined. Afterwards, several important techniques in this step are detailed.

\subsection{Background}
\label{sec:intuitive}

We firstly introduce the space labels, which comprise the basis of face membership classification. The space label of face $F$ with respect to primitive $M$, denoted as $L_F(M)$, is the relative location of $F$ with respect to $M$. Also, we use $L(M)$ to denote the space label with respect to $M$ which does not specify a face. In general, space label has four conditions: completely inside (\emph{In}), completely outside (\emph{Out}), on the boundary (\emph{On}) or not available (\emph{N/A}). In the last condition, $F$  crosses the boundary of $M$. Accordingly, it does not have a uniform label. In addition, if the normal direction of $F$ is considered, there are two derived conditions of \emph{On}: the normal points to the outside (\emph{Same}) and those to the inside (\emph{Oppo}) of $M$.

The evaluation of Boolean operations between B-reps primitives $A$ and $B$ can be converted to the problem of surface selection according to the labels \cite{kuratowski1968set}:
\begin{equation}
\label{eq:select}
\begin{split}
&A\cup^*B:\{F_A~Out~B\}\cup\{F_B~Out~A\}\cup\{F_A~Same~B\},\\
&A\cap^*B:\{F_A~In~B\}\cup\{F_B~In~A\}\cup\{F_A~Same~B\},\\
&A-^*B:\{F_A~Out~B\}\cup\{(F_B~In~A)'\}\cup\{F_A~Oppo~B\},
\end{split}
\end{equation}

\noindent where $F_X$ is the faces of $X$, and $F'$ means $F$ with an inverted normal. The stars ($^*$) after the operation notations ($\cup, \cap, -$) mean that the Boolean operations are regularized.

Now, consider computing $L_F(T)$, where $T$ is a CSG solid $T$ with $n$ primitives $\{M_i~| ~i = 1,2,3,...,n\}$.  Then there are a total of $n$ space labels for $F$: \{$L_F(M_i)~|~i = 1,2,3,...,n\}$ (or in vector form $\vec{L}_F$). If all elements in $\vec{L}_F$ are known and none of them is \emph{N/A}, $L_F(T)$ can be easily computed by traversing the whole CSG tree from bottom to top and progressively combining the space labels of the CSG nodes according to combination rules \cite{requicha1985boolean}:
\begin{equation}
\label{eq:comb}
\begin{split}
X\cup Out\Rightarrow X, ~~&X\cap Out\Rightarrow Out, \\
X\cap In~~\Rightarrow X, ~~&X\cup In\Rightarrow In, \\
X \cup X \Rightarrow X, ~~&X \cap X \Rightarrow X,\\
Same \cup Oppo\Rightarrow In, ~~&Same \cap Oppo\Rightarrow Out,
\end{split}
\end{equation}
where $X$ is an arbitrary label value. Note that the combination rules for difference operation are not shown here, because the difference operation can be converted into the intersection operation using De Morgan's transformations:
\begin{equation}
\label{eq:demo}
\begin{split}
A-B\Rightarrow A\cap B^c, ~~~~~~~&(A^c)^c\Rightarrow A,\\
(A\cap B)^c\Rightarrow A^c\cup B^c, ~~&(A\cup B)^c\Rightarrow A^c\cap B^c,
\end{split}
\end{equation}
where $X^c$ denotes the complement of $X$. If and only if $L_F(T) = Same$, $F$ lies on the boundary of the final model and has the correct normal; accordingly, $F$ belongs to the final result.

If some elements in $\vec{L}_F$ are \emph{N/A} or cannot be computed, we may be unable to compute $L_F(T)$. However, the CSG tree $T$ can be trimmed into a smaller one with fewer primitives to reduce the complexity of further evaluation. For example, assume we have a CSG with its Boolean expression $M_1\cup (M_2\cap M_3-M_4)$. Given the values of two labels $L(M_1)=Out$, $L(M_2)=In$, the expression can be rewritten as $Out\cup (In\cap M_3-M_4)$. Using the combination rules we can simplify the expression as $M_3-M_4$. Thus, any face that shares the same two labels $L(M_1)$ and $L(M_2)$ can be classified based on the trimmed CSG tree $M_3-M_4$, which has only two primitives.

\subsection{Face Classification Framework}
\label{sec:framework}

\begin{figure}
\centering
\includegraphics[width=3.5in]{multilvl}
\caption{Illustration of multi-level face grouping framework. Faces are grouped into smaller groups in each level. The trimmed CSG tree of a specific group is obtained by trimming the tree of its parent according to common labels of this group. When the group is not further subdivided, faces in this group are respectively classified based on the trimmed CSG tree of this group.}
\label{fig:hierachy}
\end{figure}

Space labels of neighboring faces are often coherent. That is, given two neighboring faces $F_1$ and $F_2$, $L_{F_1}(M)$ and $L_{F_2}(M)$ are often the same. Therefore, we can group together neighboring faces and evaluate them based on the smaller CSG tree, which is trimmed by the common labels. The trimmed tree contains fewer leaves, thus enabling faster computation. If face grouping can be efficiently performed, this strategy will greatly accelerate the process of face membership classification. Moreover, a multi-level grouping scheme (Fig. \ref{fig:hierachy}) may provide even greater capability to shared labels. By using such a scheme, faces are first grouped into first-level groups; then, faces of the same first-level groups are further divided into second-level groups, and so on. Labels are shared in each level and the original CSG tree is trimmed step by step according to common labels in each level. We developed this idea into a novel classification algorithm that uses a two-level grouping scheme. The first level is the \emph{inter-primitive} level; the second is the \emph{intra-primitive} level.

Before we begin face classification, pre-computation is conducted to make the subsequent analysis easier. The original CSG tree is converted into a \emph{positive} tree using the method by Rossignac and Voelcker \cite{rossignac1988active}. `Positive' means that the CSG tree contains no difference (-) operation. According to Equation \ref{eq:demo}, there may be some primitives with complement notation. The benefit is that we are not required to consider the difference operations in the following computation.

The classification starts from face grouping in the inter-primitive level. In this level, faces of the same primitives are simply grouped together. Then, we give a rough estimation of the common labels of the group. Suppose the group is primitive $M$. Its face labels with respect to primitives whose bounding boxes are totally outside of $M$'s bounding box must be $Out$ (for primitives with complement notations, $In$). The original CSG tree is trimmed into the first trimmed CSG tree (as short as the \emph{the first trimmed tree}) according to these common labels plus an extra label $L(M)$ whose value is obviously $Same$.

After that, the inter-primitive groups are further divided into smaller intra-primitive groups. To explain how to perform intra-primitive grouping, we firstly introduce an essential concept---intersection primitive (IP). The IPs of a face $F$ is defined as the primitives that $F$ intersects, excluding the primitive to which $F$ belongs (denoted as $M$). Primitives (excluding $M$) that do not intersect $F$ are called non-IPs. The set that contains all IPs of $F$ is the IP-set of $F$. That set can be computed by enumerating all the primitives in the cross list and coplanar list of $F$.

Now, we require that the faces of the same intra-primitive group meet two requirements: they should be connected together and have the same IP-set. These two requirements give these faces very good properties, as discussed in Proposition 1. The labels referenced in Proposition 1 are used to trim the first trimmed tree into the second trimmed CSG tree (as short as \emph{the second trimmed tree}). For a specific intra-primitive group, the primitives of its second trimmed tree is a subset of the common IP-set of that group. This is because all space labels with respect to non-IPs are shared within the intra-primitive group and all these labels are either $In$ or $Out$. Therefore, all non-IPs must have been trimmed from the CSG tree (to be explained in Section \ref{sec:trim}).

\vspace{0.8em}
\noindent \textbf{Proposition 1.}~~~~If a group of faces are connected together and they have the same IP-set, they must share the same labels with respect to all non-IPs. These labels are either \emph{In} or \emph{Out}.

\vspace{0.6em}
\noindent \textbf{Proof.}~~~~We only prove that the conclusion stands for adjacent faces, which is a special case of connected faces. Promoting the conclusion to the general situation is obvious. Assume $M$ is a non-IP and label $L(M)$ is different for the two adjacent faces $A$ and $B$ . Then there must be faces from $M$ between $A$ and $B$ to cause the different space labels. Additionally,  because $A$ and $B$ are adjacent, at least one of $A$ and $B$ intersects the faces from primitive $M$. This contradicts the assumption that $M$ is non-IP. Therefore, $L(M)$ must be the same for $A$ and $B$. Moreover, because $M$  is non-IP, $L(M)$ is either \emph{In} or \emph{Out}.

\begin{figure}
  \centering
  \includegraphics[width=3.5in]{mesh}\\
  \caption{A snapshot during flood filling (\emph{left}), and the final grouping result (\emph{right}). Different intra-primitive groups are distinguished by different colors. From the final grouping result, the whole mesh is divided into 9 intra-primitive groups.}\label{fig:mesh}
\end{figure}

>>>>

\usepackage{enumitem}
%\usepackage[]{algorithm2e}
\usepackage[]{algpseudocode}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
\usepackage{color}
%\usepackage[numbers,sort&compress]{natbib}
%\usepackage{ulem}
\renewcommand{\arraystretch}{1.3}

\usepackage{color}
\usepackage{bm}
%\usepackage{mathrsfs}
\usepackage[mathcal]{eucal}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{wrapfig}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\floatname{algorithm}{Algorithm}

\begin{document}
\title{Efficient and Robust Plane-Based Boolean Operations on N Meshes}
\author{%Rui~Wang,~
        %Xudong~Jiang,~
        %Hongbo~Fu,~
        %Ping~Li,~
        %Bin~Sheng~and
        %Enhua~Wu
\IEEEcompsocitemizethanks{

@@

\input{tex/intro}
\input{tex/overview}
\input{tex/inter}
\input{tex/tess}
\input{tex/class}
\input{tex/exp}
